---
title: "Reproducible_Research_PA1"
author: "David DU"
date: "Saturday, May 16, 2015"
output: html_document
---

#Step 1: Loading and preprocessing the data

The data for this assignment can be downloaded from the course web site:
Data site: [Activity monitoring data](https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip)[52K].

After downloading the data, save it in the current working directory. The use read.csv() to load the data into the R program. Show the first several rows.

```{r}
stepdata<-read.csv("activity.csv",colClass=c("numeric","character","numeric"))
head(stepdata)
```
Note:
Since we have to compare the activity patterns between weekdays and weekends, we change the column "date" into data class data.
```{r}
stepdata$date <- as.Date(stepdata$date, "%Y-%m-%d")
```

#Step 2: What is mean total number of Steps taken per day?
Calculate the total number of steps taken per day, and show the first several rows.
```{r}
totalsteps<-aggregate(steps~date,data=stepdata,sum,na.rm=T)
head(totalsteps)
```
Make a histogram of the total number of steps taken each day
```{r}
hist(totalsteps$steps,main="Total steps per day",xlab="day",col="blue")
```

Calculate and report the mean and median of the total number of steps taken per day
Here I use two ways to get the same result.
Method 1:
```{r}
summary(totalsteps$steps)
```
Method 2:
```{r}
round(mean(totalsteps$steps),0)
```
```{r}
median(totalsteps$steps)
```
*However, I don't know why the results are different.*

#Step 3: What is the average daily activity pattern?
Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
tsplot<-tapply(stepdata$steps,stepdata$interval,mean,na.rm=T)
head(tsplot)
```
```{r}
plot(row.names(tsplot),tsplot,type="l",xlab = "5-min interval",ylab = "Average across all Days",main = "Average number of steps taken", col = "blue")
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
max_interval <- which.max(tsplot)
names(max_interval)
```

#Step 4: Imputing missing values
Note that there are a number of days/intervals where there are missing values (coded as NA). The presence of missing days may introduce bias into some calculations or summaries of the data.

```{r}
stepdata_NA <- sum(is.na(stepdata))
stepdata_NA
```

Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval,etc

Fist Na replaced by mean in 5 min interval
```{r}
StepsAverage <- aggregate(steps ~ interval, data = stepdata, FUN = mean)
fillNA <- numeric()
for (i in 1:nrow(stepdata)) {
    obs <- stepdata[i, ]
    if (is.na(obs$steps)) {
        steps <- subset(StepsAverage, interval == obs$interval)$steps
    } else {
        steps <- obs$steps
    }
    fillNA <- c(fillNA, steps)
}
```
Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
new_stepdata <- stepdata
new_stepdata$steps <- fillNA
```
Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?
```{r}
StepsTotal2 <- aggregate(steps ~ date, data = new_stepdata, sum, na.rm = TRUE)
```
The Histogram is
```{r}
hist(StepsTotal2$steps, main = "Total steps by day", xlab = "day", col = "red")
```

And the mean and median is
```{r}
mean(StepsTotal2$steps)
```
```{r}
median(StepsTotal2$steps)
```

After replacing the mean is the same but the median is a little bit different.

#Step 5: Are there differences in activity patterns between weekdays and weekends?
For this part the weekdays() function may be of some help here. Use the dataset with the filled-in missing values for this part.

Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
```{r}
day <- weekdays(stepdata$date)
daylevel <- vector()
for (i in 1:nrow(stepdata)) {
    if (day[i] == "Saturday") {
        daylevel[i] <- "Weekend"
    } else if (day[i] == "Sunday") {
        daylevel[i] <- "Weekend"
    } else {
        daylevel[i] <- "Weekday"
    }
}
stepdata$daylevel <- daylevel
stepdata$daylevel <- factor(stepdata$daylevel)

stepsByDay <- aggregate(steps ~ interval + daylevel, data = stepdata, mean)
names(stepsByDay) <- c("interval", "daylevel", "steps")
```
Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).
```{r}
library(lattice)
xyplot(steps ~ interval | daylevel, stepsByDay, type = "l", layout = c(1, 2), 
    xlab = "Interval", ylab = "Number of steps")
```
